---
title: "p8105_hw2_ha2546"
author: "Hana Akbarnejad"
date: "10/04/2019"
output: github_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
  )

library(tidyverse)
library(dplyr)
library(readxl)

```

# Problem 1
## Part 1
In this problem, we work with _Mr. Trash Wheel_ data. In the first part of this problem, I tidied the _Trash Wheel_ sheet of this data, and removed columns without values and rows with NA's. Then I rounded the number of sports balls and converted the result to an integer variable.

```{r Mr.Trash Wheel}

trash_data = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-x15) %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    as.integer(sports_balls)
  )

```

## Part 2
In this part, I cleaned the precipitation data from 2017 and 2018. I omitted rows without precipitation data and also added year columns to these datasets. Then I stacked these datasets on top of each other to create a single _prcpt_data_, and converted _month_ to a character variable (English abbreviation).

```{r precipitation}

prcpt2017_data = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    year = "2017"
  )

prcpt2018_data = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(
    year = "2018"
  )

prcpt_data =
  bind_rows(prcpt2017_data, prcpt2018_data) %>%
  mutate(
    month = month.abb[month]
  )

```

In Mr. Trash Wheel data, we have `r nrow(trash_data)` observations, and `r ncol(trash_data)` of variables such as _dumpster_, _weight_ and _volume_ of trash for each dumpster, and the number of _plastic bottles_, _cigarette butts_, etc. In resulting precipitation dataset which we obtained from combining precipitation data of 2017 and 2018, we have `r nrow(prcpt_data)` total precipitation observations of variables _month_, _year_, and _total_ precipitation. According to this dataset, the total precipitation in 2018 is `r sum(pull(prcpt2018_data, total))` inches and the median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trash_data, year == 2017), sports_balls))`.

# Problem 2
## Part 1

Problem two was dealing with _FiveThirtyEight_ data which was composed of different .csv files. First, I started cleaning the data in _pols_month.csv_ file by cleaning the names and seperating date variable( _mon_ ) into _year_, _month_, and _day_ variables and then converted them to integers. I then transformed month's format to English abbreviation format. On the next step, I made a new column named _president_ which will include information from _prez_gop_ and _prez_dem_ columns, I just recoded the information in _prez_gop_ column and removed these two columns, as well as the variable _day_ that was mentioned in the problem.

Note: I observed two "2" values in _prez_gop_ while the values in this column should be either either "0" or "1". So I just assumed these values are supposed to be "1" and before recoding to the president column, I replaced them with "1".

```{r prob2_pols_month}

pols_month_data = read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    month = month.abb[(month)],
    prez_gop = replace(prez_gop, prez_gop == "2", "1"),
    president = recode(prez_gop, "1" = "gop" , "0" = "dem"),
  ) %>% 
   select(year, month, president, everything(), -starts_with("prez"), -day)

```

## Part 2

Second, I cleaned the _snp.csv_. The process for cleaning this part is very similar to part 1, the only thing different was that because of how the data is structured, I also had to alter _year_ and _month_ variables' order and arrange them as my leading variables. I did not select _day_ variable in this part as well because of consistency with previous part. Then I turned _month_ from numeric format to English abbreviation format.

```{r prob2_snp}

snp_data = read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), convert = TRUE) %>%
  select(year, month, close) %>% 
  mutate(
    month = month.abb[(month)]
  )

```

## Part 3

In the third part, I tidied up the _unemplyment_data_. This dataset is composed of 1 column for year, and 12 columns for different months of a year. To tidy this dataset, I had to transform it from wide format to long format, using unemployment rates as my values and created a new variable _month_ instead of having each month as a variable. I then cleaned the names of my columns using janitor. Deoing so makes the dataset consitent with previous datasets and this will enable us to join them on the next step.

```{r prob2_unemployment}

unemployment_data = read_csv("data/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemp_prcnt"
  ) %>% 
    janitor::clean_names()            

```

## Part 4

On the final step for this problem, I first left-joined the _snp_data_ to _pols_month_data_, using keys _year_ and _month_, and then left-joined the _unemployment_data_ to the result.

```{r join_datasets}

join_pols_snp_unemp = left_join(pols_month_data, snp_data, by = c("year", "month")) %>% 
  left_join(unemployment_data, by = c("year", "month")) 

```

## Results

In problem 2, we were dealing with _pols_month_, _snp_, and _unemployment_ datasets of the _FiveThirtyEight_ data.

After tidying these datasets, _pols_month_ dataset's dimension is `r dim(pols_month_data)` which means it has `r ncol(pols_month_data)` variables including _year_ and _month_, _president_ which shows whether the president was republican or democrat, and the number of governors ( _gov_ ), the number of senators ( _sen_ ), and the number of representatives ( _rep_ ) for both republican and democratic parties on associated dates. The dataset contains `r nrow(pols_month_data)` observations for these variables, and the range of years investigated is `r max(pull(pols_month_data, year)) - min(pull(pols_month_data, year))` years (between `r range(pull(pols_month_data, year))`).

The dimension of the resulting dataframe _snp_data_ is `r dim(snp_data)` which means it has `r nrow(snp_data)` observations of `r ncol(snp_data)`variables including _year_, _month_, and _close_ (the closing values of the S&P stock index). The data recorded in this dataset is associated with `r max(pull(snp_data, year)) - min(pull(snp_data, year))` years (between years `r range(pull(snp_data, year))`).

After cleaning up, the resulting _unemployment_ dataset is a `r dim(unemployment_data)` dimension dataset: `r nrow(unemployment_data)` observations of `r ncol(unemployment_data)` variables including _year_, _month_, and another variable called _unemp_prcnt_ which shows unemployement percentage associated with these dates. The range of years for these data equals: `r max(pull(unemployment_data, year))`-`r min(pull(unemployment_data, year))` = `r max(pull(unemployment_data, year)) - min(pull(unemployment_data, year))` years.

# Problem 3
```{r prob3_pop_names}

pop_names_data = read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    childs_first_name = str_to_title(childs_first_name),
    gender = str_to_lower(gender),
    ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "WHITE NON HISP" = "WHITE NON HISPANIC", "BLACK NON HISP" = "BLACK NON HISPANIC")
  ) %>% 
  mutate(
    ethnicity = str_to_title(ethnicity)               
  ) %>% distinct()

```

```{r prob3_Olivia_name}

olivia_pop = filter(pop_names_data, childs_first_name == "Olivia" & gender == "female") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank") %>% 
  knitr::kable()

olivia_pop
```

```{r prob3_male_pop}

male_pop = filter(pop_names_data, gender == "male" & rank == "1") %>% 
    select(year_of_birth, ethnicity, childs_first_name) %>% 
pivot_wider(
  names_from = "year_of_birth", 
  values_from = "childs_first_name") %>% 
  knitr::kable()

male_pop
```

```{r prob3_scatterplot}

pop_scatterplot = pop_names_data %>% 
  filter(gender == "male" & ethnicity == "White Non Hispanic" & year_of_birth == "2016") %>% 
  ggplot(aes(x = rank, y = count)) + geom_point()

pop_scatterplot
```

